\LARGE{\bf{Natural variation in teosinte at the domestication locus \emph{teosinte branched1} (\emph{tb1})}}\footnotemark

### 2014-08-13 ResponseToReview.tex

\documentclass[11pt]{article}
\usepackage[letterpaper, margin=1in]{geometry} %package that allows changes in margins and header/footers
\newcommand{\res}[1]{\noindent \textcolor{blue}{{#1}} \\}
\newcommand{\jri}[1]{\noindent \textcolor{red}{{#1}} \\}
\newcommand{\lev}[1]{\noindent \textcolor{green}{{#1}} \\}
\usepackage{lipsum}
\usepackage{url}
\usepackage{color}
\usepackage{multirow}
\renewcommand{\baselinestretch}{1.5}
\usepackage{natbib} % package for formatting citations
\bibliographystyle{AJB} %indicating bibliography uses the AJB.bst file for

\begin{document}

\noindent Dear Dr. Schnabel and additional editorial staff at the \emph{American Journal of Botany}, \\

We would like to thank you and two anonymous reviewers for your helpful comments and suggestions regarding our manuscript. We have addressed the majority of the issues raised in a revised version of our manuscript, but, while we feel that suggested new experiments would provide interesting results, they would require considerable extra time (many months) and are outside of the scope of this current work.  We have attached a revised version of our manuscript, including the suggested changes, and have detailed our responses in blue below. We hope that you are willing to consider our revised manuscript as a resubmission.\\\

@@

\noindent Sincerely,\\

\noindent Jeffrey Ross-Ibarra and Matthew Hufford

\section*{Response to Associate Editor:}

I thought that the neighbor-joining analysis was of little value, because it was not clear to me just how the result could be expected to provide information about Hopscotch genotype.  Why would you expect individuals with similar Hopscotch genotypes to cluster together?  Also, although you don't provide any information on the distribution of variable sites among accessions and taxa, your data set of only 48 and 40 segregating sites is almost certainly not sufficiently variable to resolve relationships among the very large number of accessions you were analyzing.  As a result, you produce mostly a huge polytomy, with a small amount of structure that appears to reflect the genetic similarities and differences already evident in Table 1.  The trees themselves are difficult for a reader to interpret, because none of the accession names at the branch tips (other than TIL) are defined in the text.\\

@@

\res{Both additional greenhouse and expression analyses would indeed be useful avenuse of further investigation.  Noentheless, these are extensive additonal experiments which we are unable to include at this time. We have however added text to the conclusions of this paper discussing their utility and the need for further work to elucidate the role of \emph{tb1} in branching architecture in teosinte: ``Future studies should examine expression levels of \emph{tb1} in teosinte with and without the \emph{Hopscotch} insertion and further characterize the effects of additional loci involved in branching architecture (e.g. \emph{gt1}, \emph{tru1}, and \emph{te1}) as well as include a more exhaustive phenotyping including all traits.  These data, in conjunction with more exhaustive phenotyping, should help reveal the ecological significance of the domesticated \emph{tb1} allele in natural populations of teosinte."}

One minor issue: in Fig. 1, it would be nice to see where the Balsas River Basin is in relation to the sampling shown.

\res{We have added this to the figure.} \jri{LV please add, don't wait for Matt.}

\section*{Response to Reviewer 2} 
Although this paper brings up an interesting question regarding the role of tb1 standing variation in teosinte, it is not able to answer it completely.  It presents results that have been published previously about the presence of hopscotch in teosinte, maize, and mexicana. \jri{Is this all the intro Reviewer 2 gave? please copy/paste review verbatim if not}

@@

13) Page 7 lines 14 - 15 "These analyses only included populations in which eight or more individuals were sampled."  How many populations did that include?\\

\res{We included 32 populations for these analyses, and have edited the text to reflect this number: ``These analyses only included populations in which eight or more individuals were sampled, totaling 32 populations."}

14) Page 7 lines 17 - 26: how many year's worth of environmental data went into your analysis? Can you hypothesize that tb1 differences can be correlated with current climactic data if the changes and subsequent selections must have happened many generations ago?\\

\res{The environmental data used from Pyh\"aj\"arvi  \emph{et al.} (2013) was one year's worth of environmental data. A previous study by Hufford \emph{et al.} (2012) has shown that environmental conditions in areas where teosinte grows have not changed significantly over the last 10,000 years. We have clarified the use of one year's worth of data in the text: ``Environmental data (One year's worth) were obtained from \url{www.worldclim.org}, the Harmonized World Soil Database (FAOHWSD) and \url{www.harvestchoice.org} and summarized by principle component analysis following Pyh\"aj\"arvi  \emph{et al.} (2013)."}

### 2014-08-13 ResponseToReview.tex

\documentclass[11pt]{article}
\usepackage[letterpaper, margin=1in]{geometry} %package that allows changes in margins and header/footers
\newcommand{\res}[1]{\noindent \textcolor{blue}{{#1}} \\}
\newcommand{\jri}[1]{\noindent \textcolor{red}{{#1}} \\}
\newcommand{\lev}[1]{\noindent \textcolor{green}{{#1}} \\}
\newcommand{\mbh}[1]{\noindent \textcolor{Dandelion}{{#1}}\\}
\usepackage{lipsum}
\usepackage{url}
\usepackage[usenames,dvipsnames]{color}
\usepackage{multirow}
\renewcommand{\baselinestretch}{1.5}
\usepackage{natbib} % package for formatting citations
\bibliographystyle{AJB} %indicating bibliography uses the AJB.bst file for

@@

\begin{document}

\noindent Dear Dr. Schnabel and additional editorial staff at the \emph{American Journal of Botany}, \\

We would like to thank you and two anonymous reviewers for your helpful comments and suggestions regarding our manuscript. We have addressed the majority of the issues raised in a revised version of our manuscript, but, while we feel that suggested new experiments would provide interesting results, they would require considerable extra time (many months) and are outside of the scope of this current work.  We have attached a revised version of our manuscript, including the suggested changes, and have detailed our responses in blue text below. We hope that you are willing to consider our revised manuscript as a resubmission.\\\

@@

\noindent Jeffrey Ross-Ibarra and Matthew Hufford

\section*{Response to Associate Editor:}

I thought that the neighbor-joining analysis was of little value, because it was not clear to me just how the result could be expected to provide information about Hopscotch genotype.  Why would you expect individuals with similar Hopscotch genotypes to cluster together?  Also, although you don't provide any information on the distribution of variable sites among accessions and taxa, your data set of only 48 and 40 segregating sites is almost certainly not sufficiently variable to resolve relationships among the very large number of accessions you were analyzing.  As a result, you produce mostly a huge polytomy, with a small amount of structure that appears to reflect the genetic similarities and differences already evident in Table 1.  The trees themselves are difficult for a reader to interpret, because none of the accession names at the branch tips (other than TIL) are defined in the text.\\

\res{We agree entirely with this assessment of the Neighbor Joining trees. 
Our intention with the NJ trees was to show that there is no clear signal of selection on the \emph{Hopscotch} genotype across sequence in the \emph{tb1} genomic region.
Presumably, if the \emph{Hopscotch} element were under strong selection within teosinte populations we would expect to see \emph{tb1} haplotype groups in an NJ tree due to linkage disequilibrium and Hill-Robertson effects.
In fact, the original discovery of selection on the \emph{Hopscotch} insertion during maize domestication was based on such a signal from linked sequence 60kb distant from the \emph{Hopscotch}. \mbh{is this true? What's the reference here?} 
We agree with the reviewer that the trees could be confusing and difficult to interpret, and the large polytomies do not provide information beyond our assertion that there is not strong haplotype structure across the \emph{tb1} genomic region that correlates with \emph{Hopscotch} genotype).
We have therefore opted to keep the NJ trees in the supplement where interested readers can view them but where they don't unnecessarily complicate our main story.
In addition, we have added explanation of the accession names.}

### 2014-08-13 ResponseToReview.tex

\noindent Jeffrey Ross-Ibarra and Matthew Hufford

\section*{Response to Associate Editor:}

I thought that the neighbor-joining analysis was of little value, because it was not clear to me just how the result could be expected to provide information about Hopscotch genotype.  Why would you expect individuals with similar Hopscotch genotypes to cluster together?  Also, although you don't provide any information on the distribution of variable sites among accessions and taxa, your data set of only 48 and 40 segregating sites is almost certainly not sufficiently variable to resolve relationships among the very large number of accessions you were analyzing.  As a result, you produce mostly a huge polytomy, with a small amount of structure that appears to reflect the genetic similarities and differences already evident in Table 1.  The trees themselves are difficult for a reader to interpret, because none of the accession names at the branch tips (other than TIL) are defined in the text.\\

\res{We agree entirely with this assessment of the Neighbor Joining trees. 
Our intention with the NJ trees was to show that there is no clear signal of selection on the \emph{Hopscotch} genotype across sequence in the \emph{tb1} genomic region.
Presumably, if the \emph{Hopscotch} element were under strong selection within teosinte populations we would expect to see \emph{tb1} haplotype groups in an NJ tree due to linkage disequilibrium and Hill-Robertson effects.
In fact, the original discovery of selection on the \emph{Hopscotch} insertion during maize domestication was based on such a signal from linked sequence 60kb distant from the \emph{Hopscotch}. \mbh{CITE} 
We agree with the reviewer that the trees could be confusing and difficult to interpret, and the large polytomies do not provide information beyond our assertion that there is not strong haplotype structure across the \emph{tb1} genomic region that correlates with \emph{Hopscotch} genotype).
We have therefore opted to keep the NJ trees in the supplement where interested readers can view them but where they don't unnecessarily complicate our main story.
In addition, we have added explanation of the accession names.}

\jri{need to add explanation of accession names in separate table or in figure legend.} \mbh{I'd say add this to the legend}

Phenotyping experiments - These are curious experiments.  First, why did you not generate a population of individuals with known genotypes and then test for the effect of genotype on phenotype?  Why take a random sample and just hope that you will get the necessary number of individuals of each genotype?  How can you randomize the growing conditions with respect to genotype, if you don't know genotypes ahead of time but instead determine them after all the growing measurements are taken?\\

\res{We agree that a planned grow-out with previously determined genotypes could be useful, but our interests lie in the effects of the \emph{Hopscotch} in the genetic background of individuals sampled from a naturally-occurring population.
It has previously been shown that \emph{Hopscotch} has a phenotypic effect in a controlled isogenic background, \mbh{CITE} and our pilot studies with a small sample of teosinte inbred lines also revealed a minor effect.
As leaf tissue can only be sampled from reasonably large seedligns and genotyping takes some time, we initially decided against trying to grow up many extra seedlings and genotype.  This is, however, the approach we used for our second-growout to even sample sizes. \jri{true?} Note that because plants were situated in the greenhouse before genotyping, genotypes were effectively randomize with respect to position in the greenhouse.}

### 2014-08-14 ResponseToReview.tex

\documentclass[11pt]{article}
\usepackage[letterpaper, margin=1in]{geometry} %package that allows changes in margins and header/footers
\newcommand{\res}[1]{\noindent \textcolor{blue}{{#1}} \\}
\newcommand{\jri}[1]{\noindent \textcolor{red}{{#1}} \\}
\newcommand{\lev}[1]{\noindent \textcolor{green}{{#1}} \\}
\newcommand{\mbh}[1]{\noindent \textcolor{Dandelion}{{#1}}\\}
\usepackage{lipsum}
\usepackage{url}
\usepackage[usenames,dvipsnames]{color}
\usepackage{multirow}
\renewcommand{\baselinestretch}{1.5}
\usepackage{natbib} % package for formatting citations
\bibliographystyle{AJB} %indicating bibliography uses the AJB.bst file for

\begin{document}

\noindent Dear Dr. Schnabel and additional editorial staff at the \emph{American Journal of Botany}, \\

We would like to thank you and two anonymous reviewers for your helpful comments and suggestions regarding our manuscript.  In the attached revision, we have addressed the majority of issues raised, but, while we feel that suggested new experiments would provide interesting results, they would require considerable extra time (many months) and are outside of the scope of this current work.  In addition to our revision, we include a detailed response to reviewers below. We hope that you are willing to consider our revised manuscript as a resubmission.\\\

@@

\noindent Jeffrey Ross-Ibarra and Matthew Hufford

\section*{Response to Associate Editor:}

I thought that the neighbor-joining analysis was of little value, because it was not clear to me just how the result could be expected to provide information about Hopscotch genotype.  Why would you expect individuals with similar Hopscotch genotypes to cluster together?  Also, although you don't provide any information on the distribution of variable sites among accessions and taxa, your data set of only 48 and 40 segregating sites is almost certainly not sufficiently variable to resolve relationships among the very large number of accessions you were analyzing.  As a result, you produce mostly a huge polytomy, with a small amount of structure that appears to reflect the genetic similarities and differences already evident in Table 1.  The trees themselves are difficult for a reader to interpret, because none of the accession names at the branch tips (other than TIL) are defined in the text.\\

\res{We agree entirely with this assessment of the Neighbor Joining trees. 
Our intention with the NJ trees was to show that there is no clear signal of selection on the \emph{Hopscotch} genotype across sequence in the \emph{tb1} genomic region.
Presumably, if the \emph{Hopscotch} element were under strong selection within teosinte populations we would expect to see \emph{tb1} haplotype groups in an NJ tree due to linkage disequilibrium and Hill-Robertson effects.
In fact, the original discovery of selection on the \emph{Hopscotch} insertion during maize domestication was based on such a signal from linked sequence 60kb distant from the \emph{Hopscotch}. \mbh{CITE} 
We agree with the reviewer that the trees could be confusing and difficult to interpret, and the large polytomies do not provide information beyond our assertion that there is not strong haplotype structure broadly in the \emph{tb1} genomic region that correlates with the \emph{Hopscotch} genotype.
We have therefore opted to keep the NJ trees in the supplement where interested readers can view them but where they don't unnecessarily complicate our main story.
In addition, we have added explanation of the accession names.}

\jri{need to add explanation of accession names in separate table or in figure legend.} \mbh{I'd add this to the legend}

@@

Phenotyping experiments - These are curious experiments.  First, why did you not generate a population of individuals with known genotypes and then test for the effect of genotype on phenotype?  Why take a random sample and just hope that you will get the necessary number of individuals of each genotype?  How can you randomize the growing conditions with respect to genotype, if you don't know genotypes ahead of time but instead determine them after all the growing measurements are taken?\\

\res{We agree that a planned grow-out with previously determined genotypes could be useful.  
However, prior work has already shown that \emph{Hopscotch} has a phenotypic effect in isogenic and inbred backgrounds, \mbh{CITE} and our interests lie in the effects of the \emph{Hopscotch} in genetic backgrounds occurring within nature.
Moreover, our sampling of individuals from this population for our grow out was not random.  
We initially genotyped a single individual from each sampling site within the population and then targeted our sample for the grow out to include a mixture of individuals from both \emph{Hopscotch} positive and negative sites.  
This procedure resulted in a high frequency of both genotypes.
We have clarified this methodology on page \mbh{X}, lines \mbh{XX}.
Finally, plants were situated randomly in the greenhouse prior to genotyping.
Regardless of whether genotype eventually became known, this still represents a randomized experimental design.}

@@

Why are sample sizes not reported?  You provide total samples sizes (eg, 206 or 216 for Phenotyping 1, depending on whether you read the Methods or the Results), but not sample sizes for each genotypic class.\\

\res{We have reported sample size information for all three genotypes on page \mbh{X} lines \mbh{XX}.} \jri{I didn't see sample sizes for each of the genotpyes. are they there? if not, please add.}

It was not clear to me what ``to detect the observed effect" means in lines 12-13 of p. 9.  Your Phenotyping 1 results show no effect of genotype on tillering.  Thus, it is not clear what effect you are referring to.\\

\res{We have clarified this in the text: ``Based on these initial data, we conducted a \emph{post hoc} power analysis using effect size data for \emph{tb1}-associated QTL from (Briggs \emph{et al.} 2007), which indicated that a minimum of 71 individuals in each genotypic class would be needed to detect the suggested effect of the \emph{Hopscotch} on tillering index from Briggs \emph{et al.} 2007."}
\mbh{I am concerned that originally our manuscript stated we conducted a power analysis based on our day 40 results (which is how I remembered the power analysis being done) and now it has been revised to say it was based on effect sizes in Briggs.  These approaches seem mutually exclusive.  Which did we use and why was this changed?}

### 2014-08-14 ResponseToReview.tex

\res{We agree entirely with this assessment of the Neighbor Joining trees. 
Our intention with the NJ trees was to show that there is no clear signal of selection on the \emph{Hopscotch} genotype across sequence in the \emph{tb1} genomic region.
Presumably, if the \emph{Hopscotch} element were under strong selection within teosinte populations we would expect to see \emph{tb1} haplotype groups in an NJ tree due to linkage disequilibrium and Hill-Robertson effects.
In fact, the original discovery of selection on the \emph{Hopscotch} insertion during maize domestication was based on such a signal from linked sequence 60kb distant from the \emph{Hopscotch}. \mbh{CITE} 
We agree with the reviewer that the trees could be confusing and difficult to interpret, and the large polytomies do not provide information beyond our assertion that there is not strong haplotype structure broadly in the \emph{tb1} genomic region that correlates with the \emph{Hopscotch} genotype.
We have therefore opted to keep the NJ trees in the supplement where interested readers can view them but where they don't unnecessarily complicate our main story.
In addition, we have added explanation of the accession names.}

\jri{need to add explanation of accession names in separate table or in figure legend.} \mbh{I'd add this to the legend}

Phenotyping experiments - These are curious experiments.  First, why did you not generate a population of individuals with known genotypes and then test for the effect of genotype on phenotype?  Why take a random sample and just hope that you will get the necessary number of individuals of each genotype?  How can you randomize the growing conditions with respect to genotype, if you don't know genotypes ahead of time but instead determine them after all the growing measurements are taken?\\

\res{We agree that a planned grow-out with previously determined genotypes could be useful.  
However, prior work has already shown that \emph{Hopscotch} has a phenotypic effect in isogenic and inbred backgrounds, \mbh{CITE} and our interests lie in the effects of the \emph{Hopscotch} in genetic backgrounds occurring within nature.
Moreover, our sampling of individuals from this population for our grow out was not random.  
We initially genotyped a single individual from each sampling site within the population and then targeted our sample for the grow out to include a mixture of individuals from both \emph{Hopscotch} positive and negative sites.  
This procedure resulted in a high frequency of both genotypes.
We have clarified this methodology on page \mbh{X}, lines \mbh{XX}.
Finally, plants were situated randomly in the greenhouse prior to genotyping.
Regardless of the fact that genotypes were eventually determined, this still represents a randomized experimental design.}

@@

Why are sample sizes not reported?  You provide total samples sizes (eg, 206 or 216 for Phenotyping 1, depending on whether you read the Methods or the Results), but not sample sizes for each genotypic class.\\

\res{We have reported sample size information for all three genotypes on page \mbh{X} lines \mbh{XX}.} \jri{I didn't see sample sizes for each of the genotpyes. are they there? if not, please add.}

It was not clear to me what ``to detect the observed effect" means in lines 12-13 of p. 9.  Your Phenotyping 1 results show no effect of genotype on tillering.  Thus, it is not clear what effect you are referring to.\\

\res{We have clarified this in the text: ``Based on these initial data, we conducted a \emph{post hoc} power analysis using effect size data for \emph{tb1}-associated QTL from (Briggs \emph{et al.} 2007), which indicated that a minimum of 71 individuals in each genotypic class would be needed to detect the suggested effect of the \emph{Hopscotch} on tillering index from Briggs \emph{et al.} 2007."}
\mbh{The previous edits here introduced a fair amount of confusion.  Our submitted version indicated we used Day 40 data from phenotyping 1 for the power analysis to plan phenotyping 2, not Briggs. Jeff and I chatted for a long while and went through previous drafts and our understanding is that Day 40 data were used to plan phenotyping 2 and then a post hoc analysis was done with the Briggs data.  I will edit the text to reflect this.  Please be very careful not to make edits that inaccurately describe our methods.  This could have caused us problems during the next round of review.}

### 2014-08-14 ResponseToReview.tex

\res{We agree entirely with this assessment of the Neighbor Joining trees. 
Our intention with the NJ trees was to show that there is no clear signal of selection on the \emph{Hopscotch} genotype across sequence in the \emph{tb1} genomic region.
Presumably, if the \emph{Hopscotch} element were under strong selection within teosinte populations we would expect to see \emph{tb1} haplotype groups in an NJ tree due to linkage disequilibrium and Hill-Robertson effects.
In fact, the original discovery of selection on the \emph{Hopscotch} insertion during maize domestication was based on such a signal from linked sequence 60kb distant from the \emph{Hopscotch}. \mbh{CITE} 
We agree with the reviewer that the trees could be confusing and difficult to interpret, and the large polytomies do not provide information beyond our assertion that there is not strong haplotype structure broadly in the \emph{tb1} genomic region that correlates with the \emph{Hopscotch} genotype.
We have therefore opted to keep the NJ trees in the supplement where interested readers can view them but where they don't unnecessarily complicate our main story.
In addition, we have added explanation of the accession names.}

\jri{need to add explanation of accession names in separate table or in figure legend.} \mbh{I'd add this to the legend}

Phenotyping experiments - These are curious experiments.  First, why did you not generate a population of individuals with known genotypes and then test for the effect of genotype on phenotype?  Why take a random sample and just hope that you will get the necessary number of individuals of each genotype?  How can you randomize the growing conditions with respect to genotype, if you don't know genotypes ahead of time but instead determine them after all the growing measurements are taken?\\

\res{We agree that a planned grow-out with previously determined genotypes could be useful.  
However, prior work has already shown that \emph{Hopscotch} has a phenotypic effect in isogenic and inbred backgrounds, \mbh{CITE} and our interests lie in the effects of the \emph{Hopscotch} in genetic backgrounds occurring within nature.
Moreover, our sampling of individuals from this population for our grow out was not random.  
We initially genotyped a single individual from each sampling site within the population and then targeted our sample for the grow out to include individuals from sampling sites where individuals were homozygous or heterozygous for the \emph{Hopscotch} allele.
Our reasoning in using this approach was that we wanted a high proportion of both \emph{Hopscotch} positive and \emph{Hopscotch} negative individuals for phenotypic comparisons and, given that the \emph{Hopscotch} allele is typically rare in teosinte, we wanted to enrich our sample for this allele. 
We were overly successful in this approach and had a higher frequency of individuals with the \emph{Hopscotch} positive allele.
For the Phenotyping 2 experiment we therefore selected individuals from a mixture of \emph{Hopscotch} positive and negative sites.
Despite this modified sampling approach, we still ended up with a higher frequency of \emph{Hopscotch} positive individuals, perhaps due to the fact that this allele appears to be under positive directional selection within the San Lorenzo population.
We have clarified this sampling methodology on page \mbh{X}, lines \mbh{XX}.
Finally, plants were situated randomly in the greenhouse prior to genotyping.
Regardless of the fact that genotypes were eventually determined, this still represents a randomized experimental design.}

@@

Why are sample sizes not reported?  You provide total samples sizes (eg, 206 or 216 for Phenotyping 1, depending on whether you read the Methods or the Results), but not sample sizes for each genotypic class.\\

\res{We have reported sample size information for all three genotypes on page \mbh{X} lines \mbh{XX} (Phenotyping 1) and on page  \mbh{X} lines \mbh{XX} (Phenotyping 2).  We have also corrected our typo for sample size in Phenotyping 1 and the Results and Methods sections are now consistent.} \jri{I didn't see sample sizes for each of the genotpyes. are they there? if not, please add.}

### 2014-08-14 ResponseToReview.tex

\res{We have reported sample size information for all three genotypes on page \mbh{X} lines \mbh{XX} (Phenotyping 1) and on page  \mbh{X} lines \mbh{XX} (Phenotyping 2).  We have also corrected our typo for sample size in Phenotyping 1 and the Results and Methods sections are now consistent.} \jri{I didn't see sample sizes for each of the genotpyes. are they there? if not, please add.}

It was not clear to me what ``to detect the observed effect" means in lines 12-13 of p. 9.  Your Phenotyping 1 results show no effect of genotype on tillering.  Thus, it is not clear what effect you are referring to.\\

What was the purpose of the Phenotyping 2  experiment?  You already showed no effect of genotype on phenotype using populations where you are most likely to see segregation for the insertion, so it was not clear to me why sampling more broadly, especially from populations with no evidence of the Hopscotch insertion, would be expected to improve your understanding of this relationship.\\

\res{On Day 40 of the Phenotyping 1 experiment we saw a weak positive correlation between the \emph{Hopscotch} insertion and tillering index (p=0.0848).  
Using data from Phenotyping 1, we conducted a power analysis to determine the number of individuals we would need in both homozygous genotypic classes to observe a significant result.
This analysis indicated we would need 71 individuals in both classes to detect a significant difference.
We have clarified this on page \mbh{X}, lines \mbh{X}.}

\mbh{I don't think the next paragraph is necessary}

@@

\res{Note that all the sampling for both phenotyping experiments was limited to a single large population from which we had a large number of seed collected. Seed was collected from various "sampling sites" within the contiguous population at which we had estimated Hopscotch allele frequency. For our more broad genotyping of both landrace maize, ssp. \emph{parviglumis}, and ssp. \emph{mexicana} we sampled as many populations as we could. For most of these populations the prevalence of the Hopscotch was previously unknown, and we were interested in having an idea of the overall distribution and frequency of the Hopscotch throughout these locations.}

The discussion of introgression, genetic drift, and selection in your Discussion section seems to lack coherence.  It sometimes focuses on explaining the unexpectedly high frequency of Hopscotch in present-day populations, whereas other times it seems to be addressing the presence or absence of Hopscotch in teosinte in general, and it also sometimes seems to be addressing selection on the tb1 locus.

\res{We have edited this section for clarity to better emphasize our main point that the \emph{Hopscotch} allele, which is known to produce a domesticated phenotype, is at higher than expected frequency in a subset of teosinte populations and may play an adaptive role in certain environments encountered by teosinte}

### 2014-08-20 manuscript.tex
edits to response
Removing phenotyping 2
Edits to manuscript to reflect removal of phenotyping 2 for
resubmission to AJB

\emph{Key results:} Genotyping results suggest the \emph{Hopscotch} element is at higher than expected frequency in teosinte. Analysis of linkage disequilibrium near \emph{tb1} does not support recent introgression of the \emph{Hopscotch} allele from maize into teosinte.  Population genetic signatures are consistent with selection on this locus revealing a potential ecological role for \emph{Hopscotch} in teosinte. Finally, two greenhouse experiments with teosinte do not suggest \emph{tb1} controls tillering in natural populations.
\vspace*{0.2cm}

@@

\textbf{Phenotyping of \emph{parviglumis}---}To investigate the phenotypic effects of the \emph{Hopscotch} insertion in teosinte, we conducted an initial phenotyping trial (Phenotyping 1). We germinated 250 seeds of \emph{parviglumis} collected in Jalisco state, Mexico (population San Lorenzo) \citep{Hufford2010} where the \emph{Hopscotch} is segregating at highest frequency (0.44) in our initial genotyping sample set. In order to maximize the likelihood of finding the \emph{Hopscotch} in our association population we selected seeds from sites where genotyped individuals were homozygous or heterozygous for the insertion. We chose between 10-13 seeds from each of 23 sampling sites. We treated seeds with Captan fungicide (Southern Agricultural Insecticides Inc., Palmetto, Florida, USA) and germinated them in petri dishes with filter paper. Following germination, 206 successful germinations were planted into one-gallon pots with potting soil and randomly spaced one foot apart on greenhouse benches. Plants were watered three times a day with an automatic drip containing 10-20-10 fertilizer, which was supplemented with hand watering on extremely hot and dry days.

Starting on day 15, we measured tillering index as the ratio of the sum of tiller lengths to the height of the plant \citep{Briggs2007}. Following initial measurements, we phenotyped plants for tillering index every 5 days through day 40, and then on day 50 and day 60. On day 65 we measured culm diameter between the third and fourth nodes of each plant. Following phenotyping we extracted DNA from all plants using a modified SDS extraction protocol. We genotyped individuals for the \emph{Hopscotch} insertion following the PCR protocols listed above. Based on these initial data, we conducted a \emph{post hoc} power analysis using effect size data for \emph{tb1} associated QTL from \citet{Briggs2007}, which indicated that a minimum of 71 individuals in each genotypic class would be needed to detect the suggested effect of the \emph{Hopscotch} on tillering index.

We performed a second phenotyping experiment (Phenotyping 2) in which we germinated 372 seeds of \emph{parviglumis}, choosing equally between sites previously determined to have or not have the \emph{Hopscotch} insertion. Seeds were germinated and planted on day 7 post fruit-case removal into two gallon pots. Plants were watered twice daily, alternating between fertilized and non-fertilized water. We began phenotyping successful germinations (302 plants) for tillering index on day 15 post fruit-case removal, and phenotyped every five days until day 50. At day 50 we measured culm diameter between the third and fourth nodes. We extracted DNA and genotyped plants following the same guidelines as in Phenotyping 1.

@@

\res{should we restructure this? I thought it read clearly. Key points were 1) Hop more widely spread than previously thought; 2) evidence of selection on Hop in teo}
\jri{clean up, restructure, thank editor, explain changes}

>>>>

\emph{Key results:} Genotyping results suggest the \emph{Hopscotch} element is at higher than expected frequency in teosinte. Analysis of linkage disequilibrium near \emph{tb1} does not support recent introgression of the \emph{Hopscotch} allele from maize into teosinte.  Population genetic signatures are consistent with selection on this locus revealing a potential ecological role for \emph{Hopscotch} in teosinte. Finally, a greenhouse experiment with teosinte does not suggest \emph{tb1} controls tillering in natural populations.
\vspace*{0.2cm}

@@

\textbf{Phenotyping of \emph{parviglumis}---}To investigate the phenotypic effects of the \emph{Hopscotch} insertion in teosinte we conducted a phenotyping trial in which we germinated 250 seeds of \emph{parviglumis} collected in Jalisco state, Mexico (population San Lorenzo) \citep{Hufford2010} where the \emph{Hopscotch} is segregating at highest frequency (0.44) in our initial genotyping sample set. In order to maximize the likelihood of finding the \emph{Hopscotch} in our association population we selected seeds from sites where genotyped individuals were homozygous or heterozygous for the insertion. We chose between 10-13 seeds from each of 23 sampling sites. We treated seeds with Captan fungicide (Southern Agricultural Insecticides Inc., Palmetto, Florida, USA) and germinated them in petri dishes with filter paper. Following germination, 206 successful germinations were planted into one-gallon pots with potting soil and randomly spaced one foot apart on greenhouse benches. Plants were watered three times a day with an automatic drip containing 10-20-10 fertilizer, which was supplemented with hand watering on extremely hot and dry days.

Starting on day 15, we measured tillering index as the ratio of the sum of tiller lengths to the height of the plant \citep{Briggs2007}. Following initial measurements, we phenotyped plants for tillering index every 5 days through day 40, and then on day 50 and day 60. On day 65 we measured culm diameter between the third and fourth nodes of each plant. Following phenotyping we extracted DNA from all plants using a modified SDS extraction protocol. We genotyped individuals for the \emph{Hopscotch} insertion following the PCR protocols listed above. \lev{should we take this bit about power analysis out?} Based on these initial data, we conducted a \emph{post hoc} power analysis using effect size data for \emph{tb1} associated QTL from \citet{Briggs2007}, which indicated that a minimum of 71 individuals in each genotypic class would be needed to detect the suggested effect of the \emph{Hopscotch} on tillering index.

@@

\lev{should we restructure this? I thought it read clearly. Key points were hop more widely spread than previously and evidence of selection on Hop in two}

\jri{clean up, restructure, thank editor, explain changes}
